Conversations in time: interactive visualisation to explore structured temporal data

### 2020-08-18 tsibbletalk.Rmd
nesting and crossing syntax

# Background and motivation

## Interactive visualisation systems with linking views

@@

## Tidy temporal data and workflow

@@

# Shared temporal data for coordinated views

The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble data structure on top of a `tsibble` to allow for frictionless communication between different plots for temporal data. The `as_shared_tsibble()` function provides an entry point in the flow, turning a `tsibble` to a shared data object (i.e. `SharedTsibbleData` subclassing of `SharedData` from \CRANpkg{crosstalk}) that powers data transmission between plots. The \CRANpkg{tsibbletalk} package aims to streamline interactive graphical analysis with the focus of temporal and structured linking.

As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical linking where marking one or more observations in one category will broadcast to all other observations in this category. Given time series plots, click any data point on a line, highlighting the whole line as a result. The `as_shared_tsibble()` uses `tsibble`'s `key` data values to facilitate such linking. The `spec` argument in the `as_shared_tsibble()` function takes a symbolic formula to allow users to specify structured combinations, using `/` and `*` adopted from Wilkinson notation (10.2307/2346786) for nesting and crossing expressions respectively. By default, the specification assumes crossing structure for all units.

@@

# Slicing and dicing time

@@

# Case study: monthly domestic tourist trips in Australia

>>>>

# Related works: interactive visualisation systems with linking views

@@

# Motivation: tidy temporal data and workflow

@@

# Shared temporal data for coordinated views with examples

The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble data structure on top of a `tsibble` to allow for frictionless communication between different plots for temporal data. The `as_shared_tsibble()` function provides an entry point in the integrated flow, turning a `tsibble` to a shared instance (i.e. `SharedTsibbleData` subclassing of `SharedData` from \CRANpkg{crosstalk}) that powers data transmission across multiple views. The \CRANpkg{tsibbletalk} package aims to streamline interactive graphical analysis with the focus of temporal and structured linking.

As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to categorical linking where marking one or more observations in one category will broadcast to all other observations in this category. Given time series plots, click any data point on a line, highlighting the whole line as a result. The `as_shared_tsibble()` uses `tsibble`'s `key` variables to achieve these types of linking, and the `spec` argument takes one step further in constructing hybrid linking, such as hierarchical and categorical linking. For example, each series in the `aus_retail` data corresponds to all possible combinations of the `State` and `Industry` variables. They are intrinsically crossed with each other. If one variable is nested within another, this lends itself to a hierarchical structure, like geographical hierarchy. Such collection of inter-related time series are referred to as hierarchical and grouped time series in the literature [@fpp].

To incorporate structured specifications in the `key`, a symbolic formula can be passed to the `spec` argument. Adopting Wilkinson notations for factorial models [@Wilkinson1973], the `spec` follows the `/` and `*` operators tradition to declare nesting and crossing variables respectively. The `spec` for the `aus_retail` data is therefore specified as `spec = State * Industry`, which is the default for the presence of multiple `key` variables. Using `/` indicates if there is a hierarchy in the data. `tourism_monthly`

@@

# Slicing and dicing time with examples

@@

### 2020-08-18 tsibbletalk.tex
nesting and crossing syntax

\hypertarget{background-and-motivation}{%
\section{Background and motivation}\label{background-and-motivation}}

\hypertarget{interactive-visualisation-systems-with-linking-views}{%
\subsection{Interactive visualisation systems with linking
views}\label{interactive-visualisation-systems-with-linking-views}}

@@

\hypertarget{tidy-temporal-data-and-workflow}{%
\subsection{Tidy temporal data and
workflow}\label{tidy-temporal-data-and-workflow}}

@@

\hypertarget{shared-temporal-data-for-coordinated-views}{%
\section{Shared temporal data for coordinated
views}\label{shared-temporal-data-for-coordinated-views}}

The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk}
package, introduces a shared tsibble data structure on top of a
\texttt{tsibble} to allow for frictionless communication between
different plots for temporal data. The \texttt{as\_shared\_tsibble()}
function provides an entry point in the flow, turning a \texttt{tsibble}
to a shared data object (i.e.~\texttt{SharedTsibbleData} subclassing of
\texttt{SharedData} from \CRANpkg{crosstalk}) that powers data
transmission between plots. The \CRANpkg{tsibbletalk} package aims to
streamline interactive graphical analysis with the focus of temporal and
structured linking.

As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to
categorical linking where marking one or more observations in one
category will broadcast to all other observations in this category.
Given time series plots, click any data point on a line, highlighting
the whole line as a result. The \texttt{as\_shared\_tsibble()} uses
\texttt{tsibble}'s \texttt{key} data values to facilitate such linking.
The \texttt{spec} argument in the \texttt{as\_shared\_tsibble()}
function takes a symbolic formula to allow users to specify structured
combinations, using \texttt{/} and \texttt{*} adopted from Wilkinson
notation (10.2307/2346786) for nesting and crossing expressions
respectively. By default, the specification assumes crossing structure
for all units.

@@

\hypertarget{slicing-and-dicing-time}{%
\section{Slicing and dicing time}\label{slicing-and-dicing-time}}

@@

\hypertarget{case-study-monthly-domestic-tourist-trips-in-australia}{%
\section{Case study: monthly domestic tourist trips in
Australia}\label{case-study-monthly-domestic-tourist-trips-in-australia}}

>>>>

\hypertarget{related-works-interactive-visualisation-systems-with-linking-views}{%
\section{Related works: interactive visualisation systems with linking
views}\label{related-works-interactive-visualisation-systems-with-linking-views}}

@@

\hypertarget{motivation-tidy-temporal-data-and-workflow}{%
\section{Motivation: tidy temporal data and
workflow}\label{motivation-tidy-temporal-data-and-workflow}}

@@

\hypertarget{shared-temporal-data-for-coordinated-views-with-examples}{%
\section{Shared temporal data for coordinated views with
examples}\label{shared-temporal-data-for-coordinated-views-with-examples}}

The \CRANpkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk}
package, introduces a shared tsibble data structure on top of a
\texttt{tsibble} to allow for frictionless communication between
different plots for temporal data. The \texttt{as\_shared\_tsibble()}
function provides an entry point in the integrated flow, turning a
\texttt{tsibble} to a shared instance (i.e.~\texttt{SharedTsibbleData}
subclassing of \texttt{SharedData} from \CRANpkg{crosstalk}) that powers
data transmission across multiple views. The \CRANpkg{tsibbletalk}
package aims to streamline interactive graphical analysis with the focus
of temporal and structured linking.

As opposed to one-to-one linking, \CRANpkg{tsibbletalk} defaults to
categorical linking where marking one or more observations in one
category will broadcast to all other observations in this category.
Given time series plots, click any data point on a line, highlighting
the whole line as a result. The \texttt{as\_shared\_tsibble()} uses
\texttt{tsibble}'s \texttt{key} variables to achieve these types of
linking, and the \texttt{spec} argument takes one step further in
constructing hybrid linking, such as hierarchical and categorical
linking. For example, each series in the \texttt{aus\_retail} data
corresponds to all possible combinations of the \texttt{State} and
\texttt{Industry} variables. They are intrinsically crossed with each
other. If one variable is nested within another, this lends itself to a
hierarchical structure, like geographical hierarchy. Such collection of
inter-related time series are referred to as hierarchical and grouped
time series in the literature \citep{fpp}.

To incorporate structured specifications in the \texttt{key}, a symbolic
formula can be passed to the \texttt{spec} argument. Adopting Wilkinson
notations for factorial models \citep{Wilkinson1973}, the \texttt{spec}
follows the \texttt{/} and \texttt{*} operators tradition to declare
nesting and crossing variables respectively. The \texttt{spec} for the
\texttt{aus\_retail} data is therefore specified as
\texttt{spec\ =\ State\ *\ Industry}, which is the default for the
presence of multiple \texttt{key} variables. Using \texttt{/} indicates
if there is a hierarchy in the data. \texttt{tourism\_monthly}

@@

\hypertarget{slicing-and-dicing-time-with-examples}{%
\section{Slicing and dicing time with
examples}\label{slicing-and-dicing-time-with-examples}}

@@
